Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-12-01 response_round1.tex

%%%%%%%%%%%packages added by Spencer%%%%%%%%%%%%%
\usepackage{booktabs} %for the table in the analysis section
\usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular
\usepackage{newtxmath} %for making indicator function
\usepackage{changepage} %for adjusting the margins
\usepackage{dsfont} %for getting the indicator function
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\1}[1]{\mathds{1}\left[#1\right]}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{amsmath}

\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{red} Response: #1}}

\begin{document}
\Huge
\noindent Response to reviewers
\\~\\
\normalsize
We are grateful to the three reviewers and their insightful comments and 
questions. Overall, they made a huge difference in improving the content of
our manuscript. In the most recent revision, we are particularly grateful for 
the questions on uncertainty propogation. Obviously the explanation we gave
for how uncertainty propogation wasn't clear enough, so in this iteration of 
 we reworked sections 3.7 and 3.8 to make more clear how 
uncertainty propogation was done. Specifics are found below and in the 
revised manuscript. Reviewer comments are \emph{italicized}, our responses to 
each reviewer's comments are in normal script, 
and specific changes/additions to the main manuscript are 
indented.
\\~\\

@@

\Huge
\noindent AE1
\normalsize
\\~\\

% \emph{
% I think the authors did a great job in this newly submitted version and 
% addressed most of the earlier raised comments in a nice and convincing way. 
% The paper is well written and interesting to read. 
% }
% I only have two remaining comments.

@@

\begin{enumerate}[-]

  \item \emph{Section 3.7. End of of first paragraph: ``..., but propagation of 
  uncertainty   in ILI forecasts was stil permitted in the estimating of 
  hospitalisation forecasts.``
  }
  \emph{
  Please clarify more explicitly how uncertainty was propagated between the two 
  model components. My understanding is that this was achieved by incorporating 
  not only point estimates of the parameters but also samples from the predictive 
  distribution of the ILI model. If this is correct, please state it clearly 
  already at this stage and consider adding a brief discussion of the advantages 
  and disadvantages of this approach.}
  \\~\\
  To clarify how uncertainty was propogated, we consildated sections
  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
  also added some advantages/disadvantages of our approach. The changes now
  read:
  \begin{quote}
        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
        \label{sec:implementation_posterior}

        Rather than combine the ILI and hopsitalization models into a single model,
        (1) and 
        (6) were fit separately via Bayesian posterior updating 
        in two stages and then fused together to produce hospitalization 
        forecasts. In the first stage, future ILI 
        forecasts were obtained from the posterior predictive distribution where for
        week $w$, the predictive distribution was obtained by integrating over the 
        parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
        as in (\ref{eq:ili_post}) where 
        $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
        is the density function of the posterior distribution for the model parameters
        and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
        The desired forecasts were for weeks $w + i$ 
        where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
        predicted or forecasted ILI at $i$ weeks in the future.

@@

        \begin{equation}
            \label{eq:ili_post}
            \tag{7}
            p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
            p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
            \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
            \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
            d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
        \end{equation}
        % \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
        % \spencer{Isn't that captured in $d \pi_s$?}

        \indent In the second stage, the posterior distribution
        for parameters of model (6) given observed
        hospitalization data $\textbf{H}_w$, observed ILI data, and
        denoted as 
        $p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
        was obtained.
        The posterior predictive in (\ref{eq:ili_post}) was then combined with 
        the hospitalization parameter posterior distribution
        to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
        for week $w + i$.
        Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
        parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
        That is, the first stage component model produced $K$ draws
        $(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
        and the second stage component model produced
        $K$ parameter draws 
        $(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
        $(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
        ILI and parameter posterior distribution draws was then plugged into equation
        (6) to produce

        \[
         \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
                                          \alpha_{1s}^{(k)} 
                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
                                          \alpha_{2s}^{(k)} 
                                          (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
                                          \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
                                          \epsilon_{s,w + i}^{(k)}
        \]
        where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
        $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
        from which
        the probabilistic hospitalization forecast was estimated. In practice we set
        $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
        \\~\\
        \indent Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
        values,
        which do not make sense when the distribution is meant to forecast
        hospitalizations, a nonnegative number. We attempted to alleviate this problem
        by modeling log-hospitalizations or by assuming hospitalizations followed a
        distribution truncated at 0, but these models led to poorer forecasts
        and issues drawing from the posterior distribution.
        For the forecast competition analysis in 
        section 5, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
        were 
        set to 0 to reflect realistic values of hospitalizations and comply with the 
        FluSight forecasting rules.
        \\~\\

        \indent In the two-stage scheme, the posterior distribution of 
        model (1) parameters was not directly influenced by the 
        hospitalization data and the posterior distribution of model 
        (6) parameters was not directly influenced by model 
        (1) parameters. Hence there was limited feedback between the
        two components, but propagation of uncertainty in ILI forecasts was still 
        permitted in the estimating of hospitalization forecasts.
        This two-stage modeling is remeniscent of the Bayesian cut 
        described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
        into two component models or ``modules`` may be justified for several reasons
        including to reduce time to fit the model, avoiding mixing issues of posterior
        sampling distributions, improving predictive performance or to prevent
        model misspecification in one model to influence the other 
        \cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
        lee2017rigorous, plummer2015cuts}. A drawback
        from cutting a model is potential loss of useful information. 
        In our current 
        application for example, the ILI model parameters receive no potentially 
        useful information from hospiatlization data.
        \\~\\
        \indent The decision to cut the forecast model into two component models
        was made for three major reasons.
        When fitting the fully Bayesian joint model, the time to fit a single model 
        using
        MCMC sampling and
        obtain forecasts was sometimes many hours whereas fitting the two models
        separately took a fraction of the time. When assessing convergence of model
        parameters, it was rare that posterior sampling chains for the joint model
        showed convergence even after 55,000 iterations, whereas the component model 
        fits gave no signs which raised concern for convergence. And most importantly,
        the forecasts from the component models generally outperformed those of the 
        joint model in terms of minimizing proper scoring rules.
        \\~\\
  \end{quote}
  \emph{
  \item I recommend including the statement raised by the reviewer in the first 
  submission—that WIS approximates CRPS (ideally supported with a reference)—and 
  therefore leads to essentially the same conclusions. I think the motivation for 
  using WIS instead of CRPS is well explained, especially given that WIS is the 
  primary evaluation measure used in the FluSight competition} 
  \\~\\

  The following statements were added to page 17.
  \begin{quote}

      Commonly used proper scoring rules for continuous distribution forecasts
      include the log score and the continuously
      ranked probability score (CRPS).
      \\~\\

      We note that as the number of forecast intervals $B$ increases, the WIS becomes 
      arbitrarily close to
      the CRPS. Thus, forecast scores under the CRPS and the WIS should be similar
      \cite[]{bracher2021evaluating}.
      \\~\\
\end{quote}

@@

\end{enumerate}

\Huge
\noindent R1
\normalsize
\\~\\
\noindent \emph{The authors have addressed all my comments from the first round. I only have two remaining points
on the extended empirical analysis on FluSight.}

@@

\begin{enumerate}[1.]

  \item \emph{The authors mention the difficulty of missing submissions during the FluSight season (hence the
    display chosen for Figure 10). How was this handled for Table 2? Do the averages refer to different
    sets of forecasting tasks for each model? If so the results are not fully comparable. The referenced
    paper by Cramer et al contains a (heuristic) scheme to compare scores in such a setting, which
    resembles what was done for Figure 10. This may be more appropriate than reporting raw averages.
    Models with very small submission fractions (maybe those with less than 80\%, definitely those with
    less than 50\%) in my opinion should just be removed altogether.}
    \\~\\
          The columns RLWIS and RWIS in table 2 correspond directly to the 
          values plotted in figure 10. We rearranged the table so that RLWIS and
          RWIS come first. The updated table, shown as "Table 1" is below.
          To clarify what's given in table 2 we stated the
          following.

    \begin{quote}
      Table 2 gives the numerical values for RLWIS 
      and relative WIS (RWIS) over
      all locations and weeks for each model, relative to the ASGD\_NORM2 model.
      It also includes the non relative mean WIS and LWIS for each model, that is the 
      overall mean scores for all forecasts submitted regardless of submission overlap
      with another model.
    \end{quote}

    \begin{quote}

      During the FluSight competition, most participating teams would occasionally 
      miss a forecast submission for some week, location, or horizon.
      The FluSight 
      models which are used for comparison are the
      13 non-ensemble models which submitted forecasts for at least 80\% of 
      FluSight targets during the 2023 season.

      Because forecast submissions for some targets were occassionally missed
      by participating teams, for any two models there was often a mismatch between
      targets for which forecasts were included or exlcuded. This makes a
      simple average of scores is an imperfect measure for comparison.
    \end{quote}

  \item \emph{It is uncommon to report median scores as their statistical properties are unclear. As mean score
    are also there I'm not fundamentally opposed to reporting them, but in my opinion mean scores
    alone are sufficient.}  
    \\~\\
    Thank you for this comment. Because of it, we removed the median scores
    from the table which improved the aesthetic of the table as well as simplified
    it. See the new table below.
\\~\\    
\end{enumerate}

@@

\newpage

\Huge
\noindent R2
\normalsize
\\~\\
% Forecasting Influenza Hospitaliztions Using a Bayesian Hierarchical Nonlinear 
% Model with Discrepancy
% 
% 
% This manuscript seems to be a very reasonable attempt at forecasting influenza 
% induced hospitalizations. The
% manuscript also puts the methodology in the context of the FluSight competition 
% organised by the CDC in the US.
% Apologies, I have found it hard to review as it is conceptually very hard to do 
% so in my opinion. Firstly,
% I do not find comfortable modelling the proportion, rather a natural outcome to 
% model will be the number of hospitalization using a Generalized Linear model 
% (Poisson distribution). Secondly, I find it un-intuitive to model the
% flu curve using an Asymmetric Gaussian (ASG) Function. For a statistical model 
% we focus on two aspects: the systematic part and the random part. Here, as far 
% as  I understand the systematic part is the ASG function. The random
% part has several components, which are hidden in Equation (5). My difficulty in 
% modelling using the ASG
% function lies in my inability to relate any meaningful meteorological variables 
% (which are very much drivers of the seasonal flu outbreaks). See for example 
% the paper,  ``A hierarchical Bayesian model for improving short-term forecasting 
% of hospital demand by including meteorological information`` by  SK Sahu, 
% B Baffour, PR Harper, JH Minty, C Sarran
% Journal of the Royal Statistical Society, Series A (2014).

% In spite of the above criticisms, 
\emph{I appreciate the hierarchical modeling 
approach of fitting the two models separately in components. But I wonder how 
the uncertainty has been propagated from one component model to another. 
Cutting the feedback idea is well understood but one still needs to integrate 
uncertainty to get an overall measure of the uncertainty in the forecasts. 
I did not clearly understand this issue from my reading of  the paper. I can 
suggest one way to propagate this uncertainty is by using MCMC samples 
(and methods), see for example, the paper,  ``A rigorous statistical framework 
for spatio-temporal pollution prediction and estimation of its long-term impact 
on health`` by 
D Lee, S Mukhopadhyay, A Rushworth, and SK Sahu, in Biostatistics (2017).}
\\~\\
We appreciated this comment and the included reference. We used the reference
to assist in our new explanation, and made sure to cite it. 
To clarify how uncertainty was propogated, we consildated sections
  3.7 and 3.8, reordered the main ideas, and removed redundant information. We
  also added some advantages/disadvantages of our approach. The changes now
  read:
  \begin{quote}
        \subsection*{3.7 Hospitalization Forecasts Using the Bayesian Cut}
        \label{sec:implementation_posterior}

        Rather than combine the ILI and hopsitalization models into a single model,
        (1) and 
        (6) were fit separately via Bayesian posterior updating 
        in two stages and then fused together to produce hospitalization 
        forecasts. In the first stage, future ILI 
        forecasts were obtained from the posterior predictive distribution where for
        week $w$, the predictive distribution was obtained by integrating over the 
        parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
        as in (\ref{eq:ili_post}) where 
        $p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
        is the density function of the posterior distribution for the model parameters
        and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
        The desired forecasts were for weeks $w + i$ 
        where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
        predicted or forecasted ILI at $i$ weeks in the future.

@@

        \begin{equation}
            \label{eq:ili_post}
            \tag{7}
            p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
            p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
            \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
            \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
            d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
        \end{equation}
        % \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
        % \spencer{Isn't that captured in $d \pi_s$?}

        \indent In the second stage, the posterior distribution
        for parameters of model (6) given observed
        hospitalization data $\textbf{H}_w$, observed ILI data, and
        denoted as 
        $p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
        was obtained.
        The posterior predictive in (\ref{eq:ili_post}) was then combined with 
        the hospitalization parameter posterior distribution
        to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
        for week $w + i$.
        Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
        parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
        That is, the first stage component model produced $K$ draws
        $(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
        and the second stage component model produced
        $K$ parameter draws 
        $(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
        $(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
        ILI and parameter posterior distribution draws was then plugged into equation
        (6) to produce

### 2025-12-03 manuscript_res.tex
clean up section 3.7

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network collects 
information on respiratory illness from outpatient visits to health care 
providers. Each week, over 3,400 outpatient health care providers in all 50 
US states, PR, 
DC, and the US Virgin Islands report the total number of outpatient 
visits along with the number of ILI cases. An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat.`` Prior to the 2021 season, the definition included 
``without a known cause other than influenza`` \cite[]{cdcfluview2023}. Because 
other illnesses such as COVID-19, RSV, and the common cold may induce similar 
respiratory symptoms, ILI may include patients infected with some disease 
other than influenza. To know whether or not a sick patient is infected with 
influenza requires a laboratory test.

@@

The prior distributions
selected for the ILI data model under both the SIR and ASG models 
largely follow the prior distribution 
selections in Osthus et al. \cite[]{osthus2019dynamic} 
and Ulloa \cite[]{ulloa2019} with a few exceptions where selecting
different priors improved 
identifiability, 
numerical stability, and/or we felt the adjusted prior made more sense for the 
problem. For model (\ref{eq:ili_model}), parameters which are common even when 
using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
$\sigma_{\gamma}^2$, $\sigma_{\gamma_W}^2$, and $\sigma_{\upsilon}$. For the SIR function 
$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
function $\theta_s = (\lambda_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
estimated is $\Psi = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

@@

Rather than combine the ILI and hopsitalization models into a single model,
(\ref{eq:ili_model}) and 
(\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating 
in two stages and then fused together to produce hospitalization 
forecasts. In the first stage, future ILI 
forecasts were obtained from the posterior predictive distribution where for
week $w$, the predictive distribution was obtained by integrating over the 
parameters $\pi_s$, $\kappa_s$, $\sigma^2_{\gamma}$, and $\sigma^2_{\gamma_W}$ 
as in (\ref{eq:ili_post}) where 
$p(\boldsymbol{\pi}_s, \kappa_s, \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w)$ 
is the density function of the posterior distribution for the model parameters
and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$. 
The desired forecasts were for weeks $w + i$ 
where $i$ is a positive integer. Here $\widetilde{ILI}_{s,w + i}$ represents the 
predicted or forecasted ILI at $i$ weeks in the future.

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
    p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
    \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
    d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}

In the second stage, the posterior distribution
for parameters of model (\ref{eq:final_hosp}) given observed
hospitalization data $\textbf{H}_w$, observed ILI data, and
denoted as 
$p(\boldsymbol{\alpha}_s, \sigma_{\epsilon_s} | \textbf{H}_w, \textbf{ILI}_w)$ 
was obtained.
The posterior predictive in (\ref{eq:ili_post}) was then combined with 
the hospitalization parameter posterior distribution
to produce a hospitalization forecast $\widetilde{H}_{s,w + i}$ 
for week $w + i$.
Both the ILI posterior predictive in (\ref{eq:ili_post}) and the hospitalization
parameter posterior were estimated via Markov chain Monte Carlo (MCMC) sampling.
That is, the first stage component model produced $K$ draws
$(\widetilde{ILI}_{s,w +i}^{(1)}, ..., \widetilde{ILI}_{s,w +i}^{(K)})$ 
and the second stage component model produced
$K$ parameter draws 
$(\boldsymbol{\alpha}_s^{(1)}, ..., \boldsymbol{\alpha}_s^{(K)})$ and 
$(\epsilon_{s,w}^{(1)}, ..., \epsilon_{s,w}^{(K)})$. Each of the
ILI and parameter posterior distribution draws was then plugged into equation
(\ref{eq:final_hosp}) to produce

\[
 \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
                                  \alpha_{1s}^{(k)} 
                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
                                  \alpha_{2s}^{(k)} 
                                  (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
                                  \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
                                  \epsilon_{s,w + i}^{(k)}
\]
where $k \in \{1, ..., K\}$. The hospitalization draws were then combined into
$\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
from which
the probabilistic hospitalization forecast was estimated. In practice we set
$K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.

Occasionally $\{H^{(k)}_{s,w + i}\}^K$ included a small number of negative 
values,
which do not make sense when the distribution is meant to forecast
hospitalizations, a nonnegative number. We attempted to alleviate this problem
by modeling log-hospitalizations or by assuming hospitalizations followed a
distribution truncated at 0, but these models led to poorer forecasts
and issues drawing from the posterior distribution.
For the forecast competition analysis in 
section \ref{sec:analysis}, all negative values of $\{H^{(k)}_{s,w + i}\}^K$ 
were 
set to 0 to reflect realistic values of hospitalizations and comply with the 
FluSight forecasting rules.

In the two-stage scheme, the posterior distribution of 
model (\ref{eq:ili_model}) parameters was not directly influenced by the 
hospitalization data and the posterior distribution of model 
(\ref{eq:final_hosp}) parameters was not directly influenced by model 
(\ref{eq:ili_model}) parameters. Hence there was limited feedback between the
two components, but propagation of uncertainty in ILI forecasts was still 
permitted in the estimating of hospitalization forecasts.
This two-stage modeling is remeniscent of the Bayesian cut 
described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
into two component models or ``modules`` may be justified for several reasons
including to reduce time to fit the model, avoiding mixing issues of posterior
sampling distributions, improving predictive performance or to prevent
model misspecification in one model to influence the other 
\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
lee2017rigorous, plummer2015cuts}. A drawback
from cutting a model is potential loss of useful information. 
In our current 
application for example, the ILI model parameters receive no potentially 
useful information from hospiatlization data.

@@

Both component models were fit via MCMC sampling using the 
\texttt{cmdstanr} package which was developed and is maintained by the Stan 
Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
implements Hamiltonian Monte Carlo (HMC) sampling with the No-U-turn 
sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
several diagnostic statistics for assessing the sampler.
Plots of prior and posterior distributions for select parameters from ILI and 
hospitalization models are included in the supplementary material 
\cite[]{wadsworth2025bas}.

>>>>

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network collects 
information on respiratory illness from outpatient visits to health care 
providers. Each week, over 3,400 outpatient health care providers in all 50 
US states, PR, 
DC, and the US Virgin Islands report the total number of outpatient 
visits along with the number of ILI cases. An ILI case is defined as a 
``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough 
and/or a sore throat.'' Prior to the 2021 season, the definition included 
``without a known cause other than influenza'' \cite[]{cdcfluview2023}. Because 
other illnesses such as COVID-19, RSV, and the common cold may induce similar 
respiratory symptoms, ILI may include patients infected with some disease 
other than influenza. To know whether or not a sick patient is infected with 
influenza requires a laboratory test.

@@

The prior distributions
selected for the ILI data model under both the SIR and ASG models 
largely follow the prior distribution 
selections in Osthus et al. \cite[]{osthus2019dynamic} 
and Ulloa \cite[]{ulloa2019} with a few exceptions where selecting
different priors improved 
identifiability, 
numerical stability, and/or we felt the adjusted prior made more sense for the 
problem. For model (\ref{eq:ili_model}), parameters which are common even when 
using different functions of $f_{\theta_s}(w)$ are $\kappa_s$, 
$\sigma_{\gamma}^2$, $\sigma_{\gamma_W}^2$, and $\sigma_{\upsilon}$. For the SIR function 
$\theta_s = (S_{0s}, I_{0s}, R_{0s}, \alpha_s, \rho_s)$, and for the ASG 
function $\theta_s = (\lambda_s, \eta_s, \mu_s, \sigma_{1s}^2, \sigma_{2s}^2)$. 
For the hospitalization model in (\ref{eq:final_hosp}) the parameter to be 
estimated is 
$\psi_s = (\alpha_{0s}, \alpha_{1s}, \alpha_{2s}, \phi, \sigma_{\epsilon_s}, \omega_s)$.

@@

Rather than combine the ILI and hopsitalization models into a single model,
the ILI model (\ref{eq:ili_model}) and hospitalization model
(\ref{eq:final_hosp}) were fit separately via Bayesian posterior updating 
in two stages and then fused together to produce hospitalization 
forecasts. In the first stage, future ILI 
forecasts were obtained from the ILI forecast model in (\ref{eq:ili_post}).
Here 
$p(\xi | \textbf{ILI}_w)$ 
is posterior distribution from the ILI model (\ref{eq:ili_model})
and $\textbf{ILI}_w$ represents all observed ILI data up to week $w$.
The ILI forecast model was obtained by integrating over the 
parameters 
$\xi = (\pi$, $\kappa$, $\sigma^2_{\gamma}$, $\sigma^2_{\gamma_W})$.
The desired forecasts are for weeks $w + i$ 
where $i$ is a positive integer, and $\widetilde{ILI}_{w + i}$ represents the 
predicted or forecasted ILI at $i$ weeks in the future.
Note that
in this section
the parameter subscript for season $s$ is dropped to simplify notation.

% \begin{equation}
%     \label{eq:ili_post}
%     p(\widetilde{ILI}_{s,w +i} | \textbf{ILI}_w) = \int \int \int \int 
%     p(\widetilde{ILI}_{s,w + i} | \boldsymbol{\pi}_s, \kappa_s, 
%     \sigma^2_{\gamma}, \sigma^2_{\gamma_W}) p(\boldsymbol{\pi}_s, \kappa_s, 
%     \sigma^2_{\gamma}, \sigma^2_{\gamma_W} | \textbf{ILI}_w) 
%     d\boldsymbol{\pi}_s d \kappa_s d \sigma^2_{\gamma} d \sigma^2_{\gamma_W}
% \end{equation}

\begin{equation}
    \label{eq:ili_post}
    p(\widetilde{ILI}_{w +i} | \textbf{ILI}_w) = \int  
    p(\widetilde{ILI}_{w + i} | \xi) 
    p(\xi | \textbf{ILI}_w) 
    d\xi
\end{equation}
% \jarad{It doesn't make sense to have 4 $\int$ here since $\pi_s$ is a vector.}
% \spencer{Isn't that captured in $d \pi_s$?}

In the second stage, the posterior distribution
for parameters of the hospitalization 
model (\ref{eq:final_hosp}) given observed
hospitalization data $\textbf{H}_w$ and $\textbf{ILI}_w$ was obtained.
The posterior distribution is denoted as
$p(\psi | \textbf{H}_w, \textbf{ILI}_w)$.
The ILI forecast model in (\ref{eq:ili_post}) was then combined with 
the hospitalization posterior distribution
to produce a hospitalization forecast $\widetilde{H}_{w + i}$ 
for week $w + i$.
Both the ILI forecast model in (\ref{eq:ili_post}) and the hospitalization
posterior distributions were estimated via Markov chain Monte Carlo 
(MCMC) sampling.
In the first or ILI stage, $K$ MCMC draws
$(\widetilde{ILI}_{w +i}^{(1)}, ..., \widetilde{ILI}_{w +i}^{(K)})$
from the forecast model were produced,
and in the second or hospitalization stage
$K$ draws $(\psi^{(1)}, ..., \psi^{(K)})$ 
from the hospitalization posterior 
were produced. Each of the
ILI forecast and hospitalization draws was then plugged into 
the hospitalization model
(\ref{eq:final_hosp}) to produce the approximate hospitalization forecast
distribution in the bottom line of(\ref{eq:hosp_post_pred}). 
Here $k \in \{1, ..., K\}$, 
and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.
In the analyses in sections \ref{sec:simulation2} and \ref{sec:analysis}
we set $K = 50,000$ draws.

\begin{equation}
\begin{aligned}
\label{eq:hosp_post_pred}
  &p(\widetilde{H}_{w + i} | \textbf{H}_w, \textbf{ILI}_w)\\
  &= \int p(\widetilde{H}_{w +i} | 
       \psi, \widetilde{ILI}_{w +i}, \textbf{H}_w)
       p(\psi | \textbf{H}_w, \textbf{ILI}_w) d \psi d\widetilde{ILI} \\
  &\approx \frac{1}{K} p(\widetilde{H}_{w + i}^{(k)} | \psi^{(k)},  
                      \tilde{ILI}_{w +i}^{(k)}, \textbf{H}_w, \textbf{ILI}_w) 
\end{aligned}
\end{equation}
% \[
%  \widetilde{H}_{s,w + i}^{(k)} = \alpha_{0s}^{(k)} + 
%                                   \alpha_{1s}^{(k)} 
%                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P) + 
%                                   \alpha_{2s}^{(k)} 
%                                   (\widetilde{ILI}_{s,w +i}^{(k)} \times P)^2 + 
%                                   \phi^{(k)} \widetilde{H}_{s,w + i - 1}^{(k)} + 
%                                   \epsilon_{s,w + i}^{(k)}
% \]
% where $k \in \{1, ..., K\}$. 
% The hospitalization draws were then combined into
% $\{H^{(k)}_{s,w + i}\}^K = (\widetilde{H}_{s,w + i}^{(1)}, ..., \widetilde{H}_{s,w + i}^{(K)})$,
% from which
% the probabilistic hospitalization forecast was estimated.
% In practice we set
% $K = 50,000$, and forecasts were obtained for weeks $i \in \{1, 2, 3, 4\}$.

% \[ \begin{array}{rl}
% p(\tilde{H}_w | H_w, ILI_w) 
% &=               \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | H_w, ILI_w)  d\tilde{ILI}_w \\
% &\approx  \int p(\tilde{H}_w | \tilde{ILI}_w, H_w) p (\tilde{ILI}_w | ILI_w)  d\tilde{ILI}_w
% \end{array} \]

In the two-stage scheme, the posterior distribution of the ILI
model (\ref{eq:ili_model}) parameters was not directly influenced by the 
hospitalization data and the posterior distribution of the hospitalization model 
(\ref{eq:final_hosp}) was not directly influenced by ILI model 
parameters. Hence there was limited feedback between the
two components, but the propagation of uncertainty in ILI forecasts was still 
accomplished in the estimating of hospitalization forecasts through
the forecast model (\ref{eq:hosp_post_pred}).
This two-stage modeling is reminiscent of the Bayesian cut 
described in \cite{plummer2015cuts} and \cite{nott2023bayesian}. Cutting a model
into two component models or ``modules'' may be justified for several reasons
including to reduce time to fit the model, avoiding mixing issues of posterior
sampling distributions, improving predictive performance or to prevent
model misspecification in one model to influence the other 
\cite[]{nott2023bayesian, jacob2020unbiased, jacob2017better, 
lee2017rigorous, plummer2015cuts}. A drawback
from cutting a model is potential loss of useful information. 
In our current 
application for example, the ILI model parameters receive no potentially 
useful information from hospiatlization data.

@@

Both component models were fit via MCMC sampling using the 
\texttt{cmdstanr} package which was developed and is maintained by the Stan 
Development Team \cite[]{stan2024manual} \cite[]{gabry2022stan}. Stan 
implements Hamiltonian Monte Carlo sampling with the No-U-turn 
sampler \cite[]{hoffman2014no}. The \texttt{cmdstanr} package provides 
several diagnostic statistics for assessing the sampler.
Plots of prior and posterior distributions for select parameters from ILI and 
hospitalization models are included in the supplementary material 
\cite[]{wadsworth2025bas}.
